---
title: "p8105_hw2_kb2977"
author: "Francois Ban"
date: "9/28/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

##Problem 1

Using pipline to:

* Import NYC Transit data using relative paths into a subdirectory
* Read and clean the data 
* Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
* Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r pipline}
transit_data = 
  read_csv(file = "./data/nyc_transit_data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, starts_with("station"), starts_with("route"), entrance_type, entry, vending, ada, 
         -station_location) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

Explanation of dataset:

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

Answer the following questions using these data:

There are `r nrow(distinct(transit_data, station_name, line))` distinct stations.
There are `r length(which(transit_data$ada))` stations that are ADA compliant.
There are `r mean(transit_data$)` of station entrances/exits without vending that allow entrance.
What proportion of station entrances / exits without vending allow entrance?
Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?



