---
title: "p8105_hw2_kb2977"
author: "Francois Ban"
date: "9/28/2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

##Problem 1

Using pipline to:

* Import NYC Transit data using relative paths into a subdirectory
* Read and clean the data 
* Retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 
* Convert the entry variable from character (YES vs NO) to a logical variable (the ifelse or recode function may be useful).

```{r pipline_transit}
transit_data = 
  read_csv(file = "./data/nyc_transit_data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, starts_with("station"), starts_with("route"), entrance_type, entry, vending, ada, 
         -station_location) %>%
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

Explanation of dataset:

Write a short paragraph about this dataset – explain briefly what variables the dataset contains, describe your data cleaning steps so far, and give the dimension (rows x columns) of the resulting dataset. Are these data tidy?

```{r distinct_data}
transit_data_distinct = distinct(transit_data, station_name, line, .keep_all = TRUE)
```

Short numeric summary of dataset:

There are `r nrow(distinct(transit_data_distinct, station_name, line))` distinct stations.
There are `r length(which(transit_data_distinct$ada))` stations that are ADA compliant.
There are `r round(mean(transit_data_distinct$entry & transit_data_distinct$vending == "NO"), digits = 4)` proportion of station entrances/exits without vending that allow entrance.

Below is the reformated data so that route number and route name are distinct variables.

```{r tidy}
transit_data_tidy = transit_data_distinct %>% 
  gather(key = route_number, value = route_name, route1:route11)
```

There are `r nrow(filter(transit_data_tidy, route_name == "A"))` distinct stations that serve the A train.
Of the `r nrow(filter(transit_data_tidy, route_name == "A"))` distinct stations that serve the A train, `r nrow(filter(transit_data_tidy, route_name == "A", ada == "TRUE"))` are ADA compliant.

